---
title: "Island Clinical Trial of usage of Drugs vs Focus"
subtitle: "TEAM NAME"
authors: "Ahanu Youngblood, Sike Ogieva, Tyler Fields, Corri Hickson"
date: "7/13/2022"
# DO NOT CHANGE YAML BELOW THIS LINE        
output: 
  pdf_document:
  highlight: kate
number_sections: true
classoption: fleqn
geometry: margin=0.75in
linestretch: 1.25
fontfamily: cmbright
colorlinks: true
citecolor: blue
urlcolor: blue
linkcolor: blue
header-includes: 
  # header/footer formats
  - \usepackage{fancyhdr}\pagestyle{fancy}\fancyhf{}\cfoot{\thepage}
  - \lhead{\footnotesize Incubator \copyright Brittney E. Bailey}
  # formats for headings
  - \usepackage[compact]{titlesec}
  - \titleformat{\section}[hang]{\setstretch{1}\ifnum\value{section}>0 \newpage \fi \Large}{\bfseries Part \thesection}{5mm}{}[]
  - \titleformat{\subsection}[hang]{\vspace{11pt}\setstretch{1}\large}{\large\bfseries\thesubsection}{5mm}{}[]
  - \titleformat{\subsubsection}[hang]{\bfseries}{}{0mm}{}[]
  # additional text formatting
  - \AtBeginDocument{\raggedright}
  - \AtBeginEnvironment{quote}{\large\setstretch{1}}
  - \AtBeginEnvironment{Shaded}{\setstretch{1}}
  - \definecolor{shadecolor}{RGB}{248,248,255}
---
  
```{r setup, include=FALSE}
# load packages
library(tidyverse)
library(mosaic)
library(kableExtra)
library(viridis)
library(RColorBrewer)

# set some code chunk defaults
knitr::opts_chunk$set(
  # code and plain R output display
  size = "small", tidy = FALSE, comment = "\t",
  # figure defaults
  fig.align = "center", fig.width = 3.6, fig.height = 2.4,
  # suppress warnings and messages
  warning = FALSE, message = FALSE) 

# default to black & white plot theme
theme_set(theme_classic()) 

# slower to display scientific notation; "NA"s knit as blank spaces
options(scipen = 1, knitr.kable.NA = '')
```

# Data dictionary:
* `Drug_type`: assigned drug group (Mushroom, alcohol, control)
* `Before`: focus at baseline
* `Thirty_minutes`: focus score 30 minutes after administration of drug
* `An_Hour`: Focus score 1 hour after administration of drug
* `Age`: Age of individual
* `Gender`: Gender of individual
* `Name`: name of individual
* `Village`: Village in which the individual is from
* `S_No`: Subject number for data tracking


## Read in the data
<!--
Data read into variable drug_test.
Elements (Before, Thirty_Minutes, An_Hour) combined into "Time".
Refactored time from categorical to numeric data
-->
```{r read-data}
  drug_test <- read_csv("The Islands Clinical Trial Data.csv") %>% 
  pivot_longer(cols = c(Before, Thirty_Minutes, An_Hour),
               names_to = "Time_fac",
               values_to = "Missed_questions") %>% 
  mutate(Time = case_when(Time_fac == "Before" ~ 0,
                          Time_fac == "Thirty_Minutes" ~ 30,
                          Time_fac == "An_Hour" ~ 60), 
         Gender = fct_recode(Gender, 
                             "Male" = "male",
                             "Female" = "female"),
         Drug_Type = fct_relevel(Drug_Type,
                                 "alcohol", "mushroom", "control"))

```


# Data analysis
```{r summary-statistics}
# summary statistics for continuous variables
  mean_data <- drug_test %>% 
  group_by(Drug_Type, Time) %>% 
  summarize(mean_missed = mean(Missed_questions), 
            sd_missed = sd(Missed_questions)) %>% 
  ungroup()

```
```{r, fig.width=5, fig.height= 3}
  #Line Graph of Mean Missed Questions over Time. Primary Graph.
  ggplot(mean_data, aes(x = Time,
                      y = mean_missed,
                      color = Drug_Type)) +
  geom_point()+
  geom_line(size = 1.5)+
  scale_x_continuous(breaks = c(0, 30, 60))+
  labs(x = "Time (Minutes)",
       y = "Count of Missed Letters",
       color = "Treatment",
       title = "Focus vs Substance Consumption ")+
   theme(axis.text.x=element_text(angle=45, hjust = 1),
       axis.text=element_text(size = 8),
       legend.text=element_text(size = 11))
       #axis.ticks = element_line(size = 18))
  ggsave("Mean Missed Questions over Time.png", width = 20 , height = 12, unit = "cm")
```
```{r, fig.width=5, fig.height= 3}
 #Line Graph of Missed Questions over Age of Subjects
 ggplot(drug_test, aes(x = Age,
                      y = Missed_questions,
                      color = Drug_Type))+
  geom_line(aes(group = Drug_Type)) +
  labs(x = "Age (in years)", 
       y = "Count of Missed Letters",
       color = "Drug Type",
       title = "Focus vs Age in different Substance Groups")+
  theme(axis.text.x=element_text(angle=45, hjust = 1),
       axis.text=element_text(size = 8),
       legend.text=element_text(size = 14))+
   facet_wrap( ~Drug_Type)
  ggsave("Mean Missed Questions over Age.png", width = 22 , height = 12, unit = "cm")
```
```{r, fig.width = 6, fig.height= 3}

 #Line Graph of Missed Questions over Gender of Subjects
 ggplot(drug_test, aes(x = Gender,
                      y = Missed_questions,
                      color = Drug_Type))+
  geom_boxplot(aes(group = Gender)) +
  labs(x = "Gender of Subject", 
       y = "Count of Missed Letters",
       color = "Drug Type",
       title = " Focus vs Gender in different Substance Groups")+
   facet_wrap( ~Drug_Type)+
    theme(axis.text.x=element_text(angle=45, hjust = 1),
       axis.text=element_text(size = 8),
       legend.text=element_text(size = 14))
  ggsave("Missed Questions over Gender.png", width = 24, height = 12, unit = "cm")
```
```{r , fig.width = 6, fig.height= 3}

#Line Graph of Missed Questions over Loaction of Subjects

 ggplot(drug_test, aes(x = Drug_Type,
                      y = Missed_questions,
                      color = Drug_Type))+
  geom_boxplot(aes(group = Drug_Type)) +
  labs(x = "Drug Type", 
       y = "Count of Missed Letters",
       color = "Drug Type",
       title = " Focus vs Location of Subject", )+
  facet_wrap( ~Village)+
  theme(axis.text.x=element_text(angle=45, hjust = 1),
       axis.text=element_text(size = 8),
       legend.text=element_text(size = 14))
  ggsave("Missed Questions over Location.png", width = 24, height = 12, unit = "cm")
```

```{r}
library(lme4)
#library(lmerTest)

# checking model assumptions
ggplot(drug_test, aes(x = Missed_questions, group = Drug_Type, color = Drug_Type)) +
  geom_density()

# poisson regression
time <- glmer(Missed_questions ~ factor(Time) + (1 | S_No),
                family = poisson(link = "log"), data = drug_test)
drug <- glmer(Missed_questions ~ Drug_Type + (1 | S_No),
                family = poisson(link = "log"), data = drug_test)
timedrug <- drug <- glmer(Missed_questions ~ factor(Time) + Drug_Type + (1 | S_No),
                family = poisson(link = "log"), data = drug_test)
full <- glmer(Missed_questions ~ factor(Time) + Drug_Type + 
                factor(Time)*Drug_Type + (1 | S_No),
                family = poisson(link = "log"), data = drug_test)

summary(full)
anova(timedrug, full, test = "Chisq")

ggsave("Poisson Curve.png", width = 24, height = 12, unit = "cm")
```

Because we had longitudinal data, we recruited Professor Brittney Bailey (Amherst College) to assist 
with a generalized linear mixed model to determine whether drug type affected focus. 
We found there were significant differences in the number of missed questions between treatment groups over time (p < 0.01). 

